### 2018-01-31 discussion.tex
Renamed section

% !TEX root = paper.tex
\section{Conclusion and Future Work}\label{s.conclusions}