### 2018-01-31 discussion.tex
Added conclusion and future work

% !TEX root = paper.tex
\section{Conclusion and Future Work}\label{s.conclusions}

Most urban cities in the world possess a network of ANPR cameras that is used in law enforcement and for traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we've presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections. We've identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and unplausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

Once trip data has been computed, a range of interesting applications is available. This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel. By associating a vehicle with a cluster that represents taxis, or home-to-work commuters, one can begin estimating trip mode usage across the city. However, the results presented here could benefit from extra work and further validation. Finally, future work can focus on using trip data to solve interesting research problems such as: 
\begin{enumerate*}[label=(\roman*)]
  \item real-time route recommendation using probabilistic graphical models;
  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behavior;
  \item modelling how drivers make routing choices in the presence of anomalies in the road network.
\end{enumerate*}